---
title: "52414 - lab 1"
author: "52414"
date: "10/5/2023"
output: html_document
---

```{r}
library(tidyverse) # This includes dplyr, stringr, ggplot2, .. 
library(data.table)
library(rworldmap) # world map
library(ggthemes)
library(reshape2) # melt: change data-frame format long/wide
library(e1071) # skewness and kurtosis
library(rvest)
library(corrplot)
library(moments)
library(spatstat.geom)

```

# **Solution:**

# 1.a. Loading the data via URL connection:

```{r, cache=TRUE}
democracy <- read_html("https://en.wikipedia.org/wiki/Democracy_Index")
all.tables = html_nodes(democracy, "table")  
regions <- as.data.frame(html_table(all.tables[4], fill = TRUE))
countries <- as.data.frame(html_table(all.tables[6], fill = TRUE))
components <- as.data.frame(html_table(all.tables[7], fill = TRUE))
head(regions,5)
head(countries,5)
head(components,5)
```

# 1.b. Dominant countries in the democracy index

```{r, cache=TRUE}
countries_rate <- as.data.frame(countries)
top_countries <- countries_rate %>% select(Country, X2022) %>% arrange(desc(X2022)) %>% head(5)
bottom_countries <- countries_rate %>% select(Country, X2022) %>% arrange(X2022) %>% head(5)
Average <- rowMeans(countries_rate %>% select(-c(Region,X2022.rank,Country, Regime.type)))
top_countries_avg <- countries_rate %>% mutate(Average) %>% select(Country,Average) %>% arrange(desc(Average)) %>% head(5)
bottom_countries_avg <- countries_rate %>% mutate(Average) %>% select(Country, Average) %>% arrange(Average) %>% head(5)
```

Top countries in democracy index for 2022:

```{r}
top_countries
```

By 2006 -2022 Average:

```{r}
top_countries_avg
```

Lowest countries in democracy index for 2022:

```{r}
bottom_countries
```

By 2006 -2022 Average:

```{r}
bottom_countries_avg
```

# 2.a. Box plots by regions 

```{r, cache=TRUE}
p2 <- ggplot(countries)
p2 <- p2 + geom_boxplot(aes(x=Region, y=X2022)) +theme(text = element_text(size = 6))
p2
p3 <- countries %>% select(Country,Region,X2022) %>% filter(Region == "Middle East and North Africa")
p4 <- countries %>% select(Country,Region,X2022) %>% filter(Region == "Western Europe")
out_3 <- boxplot.stats(p3$X2022)$out
out_ind_3 <- which(p3$X2022 %in% c(out_3))
out_4 <- boxplot.stats(p4$X2022)$out
out_ind_4 <- which(p4$X2022 %in% c(out_4))
p3[out_ind_3,]
p4[out_ind_4,]


```

# 2.b. Regions density plots

```{r, cache=TRUE}
p5 <- ggplot(countries,aes(X2022)) + geom_density(alpha=.2, fill= "#00BFC4") + facet_wrap(~Region)
p5 
p6 <- countries %>% 
     group_by(Region) %>%
     summarize(Mean = mean(X2022), Variance = var(X2022), Skewness = skewness(X2022), Kurtosis = kurtosis(X2022))
p6 
```

From the density plots above, we can see that they don't resemble to the normal distribution. In addition we can see how the Skewness reflect in the plots, where Middle East and North Africa and Sub-Saharan Africa have the right tail as expected with positive Skewness. Moreover, we can see that the plots of the regions, apart from North America, have a negative Skewness. North America have a low variance and have no tail at all. It has Kurtosis of 1.

# 3.a. Democracy index between 2006-2022 in selected countries and regions

```{r, cache=TRUE}
Countries_graphs <- function(df, names){
  colnames <- colnames(df)
  if (sum(is.element(names, df$Region)) > 0) { x <- "Region" }
  else x <- "Country"
  
  y <-melt(df, id.vars = x)
  
  if (x == "Region"){
    y <- y %>% filter(Region %in% names)
    y <- y%>%  mutate(year = as.Date(as.character(ISOdate(as.numeric(gsub('X', '', variable)),12,31)))) %>% 
    filter(! year< "2006-12-31") %>% mutate(val = as.double(value))
    ggplot(y, aes(x = year, y =val)) + geom_line(aes(color = Region)) + labs(title="Democracy", 
         subtitle="Between the years 2006-2022",
         caption="Source: Wikipedia",
         x="Years",
         fill="Entities") 
  }
  else{
    y <- y %>% filter(Country %in% names)
    y <- y%>%  mutate(year = as.Date(as.character(ISOdate(as.numeric(gsub('X', '', variable)),12,31)))) %>% 
    filter(! year< "2006-12-31") %>% mutate(val = as.double(value))
    ggplot(y, aes(x = year, y =val)) + geom_line(aes(color = Country))+ labs(title="Democracy Index", 
         subtitle="Between the years 2006-2022",
         caption="Source: Wikipedia",
         x="Years",
         fill="Entities") }
}
Countries_graphs(countries, c("Israel", "India", "Germany","Iran","Indonesia"))
Countries_graphs(regions,p6$Region)
```

We can see that all the regions did not had any significant changes between 2006 to 2022.

# 3.b. Clustering Countries by their democracy index

```{r, cache=TRUE}
col <- colnames(countries)

max_value <- countries %>% select( colnames(countries)[which(col <= "X2022" & col >= "X2006")]) %>% apply( 1, max)
 min_value <- countries %>% select( colnames(countries)[which(col <= "X2022" & col >= "X2006")]) %>% apply( 1, min)
 
cntr <- countries %>% mutate(change = X2022 - X2006, max_value, min_value)


cntr1 <- cntr %>% filter(change >= 1.5)
clst1 <- cntr1$Country

cntr2 <- cntr %>% filter(change <= -1.5)
clst2 <- cntr2$Country

cntr3 <- cntr %>% filter(change >= 0.75 & change <= 1.5)
clst3 <- cntr3$Country

cntr4 <- cntr %>% filter(change <= -0.75 & change >= -1.5)
clst4 <- cntr4$Country

cntr5 <- cntr %>% mutate(min_ref = X2022 - min_value) %>% filter(change <= -0.75 & min_ref  >= 0.75)
clst5 <- cntr5$Country

cntr6 <- cntr %>% mutate(max_ref = X2022 - max_value) %>% filter(change >= 0.75 & max_ref  <= -0.75)
clst6 <- cntr6$Country

cntr7 <- cntr %>% mutate(Max_min = max_value -min_value) %>% filter(Max_min <= 0.5)
clst7 <- cntr7$Country

Country <- unique(c(clst1, clst2, clst3, clst4, clst5, clst6, clst7))
clst1_7 <- as.data.frame(Country)
cntr8 <- cntr %>% anti_join(clst1_7)
clst8 <- cntr8$Country

Countries_graphs(countries, clst1)
Countries_graphs(countries, clst2)
Countries_graphs(countries, clst3)
Countries_graphs(countries, clst4)
Countries_graphs(countries, clst5)
Countries_graphs(countries, clst6)
Countries_graphs(countries, clst7)
Countries_graphs(countries, clst8)

```

**Now we will analyze all the clusters**:  
cluster 1 - we can see a pattern of improvement in democracy index.  
cluster 2 - we can see a pattern of deterioration in democracy index.  
cluster 3 - we can see a slight improvement in democracy index.  
cluster 4 - we can see a slight deterioration in democracy index.  
cluster 5 - this plot have no country at all.  
cluster 6 - we can see a pattern of improvement at first but then deterioration. a big change in the democracy index indicates an unstable regimes.  
cluster 7 - we can see many countries that all most did not change in their democracy index, because of a stable regimes.  
cluster 8 - we can see that most of the countries are in this cluster, because it is a problem to analyze and understand any pattern.

# 4. **Regimes democratic mobility**

```{r, cache=TRUE}

index_2006 <- countries$X2006
regime_2006_type <- c('Full democracy','Flawed democracy','Hybrid regime','Authoritarian')


regime_mobility <- countries %>% mutate(regime_2006 = regime_2006_type[(X2006<=10&X2006>=8)+ 2*(X2006 <8 & X2006 >= 6) + 3*(X2006 < 6 & X2006 >=4) +4*(X2006 < 4 & X2006 >=0)])
                  
regime_mobility <- regime_mobility %>% mutate(regime_pairs = paste(Regime.type,regime_2006))
rgm_prb <- matrix(c(1:16),ncol = 4,byrow = TRUE)
colnames(rgm_prb) <- paste(regime_2006_type) #Regime type in 2022
rownames(rgm_prb) <- regime_2006_type #Regime type in 2006
for (i in colnames(rgm_prb)){
  for (j in rownames(rgm_prb)){
    rgm_prb[j,i] <- length(which(regime_mobility$regime_pairs %in% paste(j,i)))/length(which(regime_mobility$Regime.type %in% i))
  }
} 
heatmap(rgm_prb)
```

We can see in this heat probability map, that as expected - the chances to move from non democratic regime to democratic one is low, and higher if the regime type is similar to the current regime type.

# 5.a. **Joining more data together**

```{r, cache=TRUE}
# Load GDP table
gdp_page <- read_html("https://en.wikipedia.org/wiki/List_of_countries_by_GDP_(PPP)_per_capita")
gdp_tables <- html_nodes(gdp_page, "table")
gdp_df <- as.data.frame(html_table(gdp_tables[2], fill = TRUE))
colnames(gdp_df) <- paste(colnames(gdp_df), "-", gdp_df[1,])
colnames(gdp_df)[1] <- "Country"
gdp_df$Country <- gsub("\\\u202F\\*", "",gdp_df$Country)
gdp_df$`CIA.8..9..10. - Estimate` <- as.double(gsub(",", "", gdp_df$`CIA.8..9..10. - Estimate`))

# Load population size table
population_page <- read_html("https://en.wikipedia.org/wiki/List_of_countries_and_dependencies_by_population")
population_table <- html_table(html_nodes(population_page, "table")[[2]])
population_df <- data.frame(population_table)
colnames(population_df) <- population_df[1,]
colnames(population_df)[2] <- "Country"
population_df$Numbers <- as.double(gsub(",","", population_df$Numbers))

# Load incarceration rates table
incarceration_page <- read_html("https://en.wikipedia.org/wiki/List_of_countries_by_incarceration_rate")
incarceration_table <- html_table(html_nodes(incarceration_page, "table")[[2]])
incarceration_df <- data.frame(incarceration_table)
colnames(incarceration_df)[1] <- "Country"
incarceration_df$Country <- gsub("\\s\\[Note]", "", incarceration_df$Country)
incarceration_df$Country <- gsub("\\\u202F\\*", "",incarceration_df$Country)
incarceration_df$Rate.per.100.000..3. <- as.double(gsub(",", "", incarceration_df$Rate.per.100.000..3.))

# Load area table
area_page <- read_html("https://en.wikipedia.org/wiki/List_of_countries_and_dependencies_by_area")
area_table <- html_table(html_nodes(area_page, "table")[[2]])
area_df <- data.frame(area_table)
colnames(area_df)[2] <- "Country"
area_df$Landin.km2..mi2.<- gsub("\\s\\(.*", "", area_df$Landin.km2..mi2.)
area_df$Landin.km2..mi2.<- as.double(gsub(",", "", area_df$Landin.km2..mi2.))


# Join tables using country names
joined_table <- countries %>%
  full_join(gdp_df, by = c("Country" = "Country")) %>%
  full_join(population_df, by = c("Country" = "Country")) %>%
  full_join(incarceration_df, by = c("Country" = "Country")) %>%
  full_join(area_df, by = c("Country" = "Country"))

# Display top five rows of the joined table
head(joined_table, 5)
```

# 5.b. **Simple linear regression**

```{r, cache=TRUE}
dem_gdp <- joined_table %>% select(dem_index = X2022, cia_gdp = `CIA.8..9..10. - Estimate`, incarceration = Rate.per.100.000..3.) 
dem_gdp_lmodel <- lm(cia_gdp ~ dem_index, data = dem_gdp)  #Fitting a linear model
summary(dem_gdp_lmodel)

plot(dem_gdp$dem_index, dem_gdp$cia_gdp, pch = 16, main = "GDP PLOTTED AGAINST DEMOCRACY INDEX", xlab = "DEMOCRACY INDEX", ylab = "GDP", col = "blue") #plotting gdp and democracy index
abline(dem_gdp_lmodel, col = "purple", pch = 30, cex = 2)

dem_incar_lmodel <- lm(incarceration ~ dem_index, data = dem_gdp)
summary(dem_incar_lmodel)

plot(x = dem_gdp$dem_index, y = dem_gdp$incarceration, pch = 16, main = "INCARCERATION PLOTTED AGAINST DEMOCRACY INDEX", xlab = "DEMOCRACY INDEX", ylab = "INCARCERATION (per 100,000)", col = "red")
abline(dem_incar_lmodel, col = "orange", pch = 30, cex = 2)

```

The first plot of linear regression (the blue dots) - GDP explained by democracy index. We can see some linear connection between the variables and we have few high GDP with respectively low democracy index. The second plot (the red dots) - Incarceration explained with democracy index. We can see there is no correlation between this two.

# 6.a. **Empirical Cumulative Distribution Functions**

```{r, cache=TRUE}
X <- gdp_df$`CIA.8..9..10. - Estimate`
x_ecdf <- ecdf(X)
plot(x_ecdf,main = "GDP (PPP) per capita of a randomly selected country", xlab = "GDP", ylab = "F(X)")
```

We can see the the ECDF of GDP per country. Approximately 90% of the countries have less than 50,000$ GDP per capita. 

#6.b.

```{r, cache=TRUE}
Y_data <- joined_table %>% select(Country, gdp = `CIA.8..9..10. - Estimate`, population = Numbers) %>% filter(! is.na(gdp) & ! is.na(population)) %>% mutate(weighted_pop = population/sum(population, na.rm = TRUE))
Y_weighted_pop <- Y_data$weighted_pop
Y_gdp <- Y_data$gdp
Y_ecdf <- ewcdf(Y_gdp, Y_weighted_pop )
plot(Y_ecdf,main = "GDP (PPP) per capita of a randomly selected person", xlab = "GDP", ylab = "F(Y)")
```

We can see here the differences between X plot(6.a) to Y plot (6.b), whereas Y reflects more accurately the GDP around the world - 90% of people are below 30,000 GDP per year, very low productivity.

# 6.c.

```{r, cache=TRUE}
Z_data <- joined_table %>% select(Country, area = Landin.km2..mi2., gdp = `CIA.8..9..10. - Estimate`) %>% filter(! is.na(area) & ! is.na(gdp)) %>% mutate(weighted_area = area / sum(area, na.rm = TRUE))
Z_weighted_area <- Z_data$weighted_area
Z_gdp <- Z_data$gdp
Z_ecdf <- ewcdf(Z_gdp, Z_weighted_area)
plot(Z_ecdf,main = "GDP (PPP) per capita of a randomly selected area", xlab = "GDP", ylab = "F(Z)")

```

This plot similar to X plot (6.a) more then Y plot because it is plot of the GDP around regions and not peoples, it is more close to countries.

# 7. **Mapping democarcy index**

```{r, cache=TRUE}
country_numeric <- apply(countries[,5:19], 2, as.numeric)
country_slim_lst <- data.table(select(countries, X2022.rank, Country, X2022))
tmp_heat <-rowMeans(country_numeric[,-1:-4]) 
country_slim_lst <- mutate(.data = country_slim_lst, avg = tmp_heat)
countries_avg_rating = arrange(country_slim_lst, Country)

#we need to replace Eswatini with Swaziland
countries_avg_rating$Country = gsub("Eswatini", "Swaziland", countries_avg_rating$Country)

#again here we need to replace North Macedonia with Macedonia
countries_avg_rating$Country = gsub("North Macedonia", "Macedonia", countries_avg_rating$Country)

#using the join function to join all the tables:
heat_map_dat = joinCountryData2Map(countries_avg_rating, joinCode="NAME", nameJoinColumn="Country", verbose = TRUE)

#the white countrys are those who does not have an index of democracy
mapCountryData(heat_map_dat, nameColumnToPlot = "avg", mapTitle = "Average democracy index by Country", addLegend = TRUE)

#now we will repeat this function to show the differences between 2006 and 2022:

countries$Country = gsub("Eswatini", "Swaziland", countries$Country)
countries$Country = gsub("North Macedonia", "Macedonia", countries$Country)
democracy_index = countries %>% select(Country, X2022, X2006) %>%
  mutate(Difference = (X2022 - X2006)) %>% arrange(Country)

dimocracymaps_heat = joinCountryData2Map(democracy_index, joinCode="NAME", nameJoinColumn="Country", verbose = TRUE)

mapCountryData(dimocracymaps_heat, nameColumnToPlot = "Difference", mapTitle = "The difference index between 2022 to 2006 of democracy", addLegend = TRUE)

```

At the first heat map of the world we can see that the western world is more democratic then the eastern side (except Australia and most of Europe). The second heat map show the difference index between 2022 to 2006. we can see that the south world became more democratic.

# 8.a. **Democracy components**

```{r, cache=TRUE}

#we will change the columns names
colnames(components) <- c("Rank", "chage_in_rank", "Country",  "Regime_type", "Overall_score", "chage_in_score", "Electoral_process_and_pluralism", "Functioning_of_government", "Political_participation", "Political_culture", "Civil_liberties")

# Removing all the "Na" cells
components <- components[complete.cases(components), ]

#Using the Union for all columns, as numeric
components_numeric <- c("Electoral_process_and_pluralism", "Functioning_of_government", "Political_participation", "Political_culture", "Civil_liberties")

# Using the as.numeric function to verify its all numeric
components[, components_numeric] <- sapply(components[, components_numeric], as.numeric)

#Using the merge function to merge all the columns
total_data <- joined_table %>% full_join(components, by = c("Country" = "Country"))

#Display the top 5 rows:
head(total_data, 5)

#Extracting data for correlation matrix


cor_data <- total_data %>% select(Electoral_process_and_pluralism, Functioning_of_government, Political_participation, Political_culture, Civil_liberties) %>% filter(! is.na(Civil_liberties))

cor_matrix <- round(cor(cor_data),2)
#correlation heatmap
heatmap(cor_matrix)
```

This is the heat map of the correlation of the components of democracy.

# 8.b. **Multiple linear regression**

```{r, cache=TRUE}
outlier_table <- total_data %>% select( Country,cia_gdp = `CIA.8..9..10. - Estimate` ,Electoral_process_and_pluralism,Functioning_of_government,Political_participation, Political_culture,Civil_liberties) %>% filter(! is.na(cia_gdp) & !is.na(Electoral_process_and_pluralism) & !is.na(Functioning_of_government) & !is.na(Political_participation) & !is.na(Political_culture) & !is.na(Civil_liberties))

democracy_gdp_model <- lm(formula = cia_gdp ~ Electoral_process_and_pluralism + Functioning_of_government + Political_participation + Political_culture + Civil_liberties, data = outlier_table)
summary(democracy_gdp_model)

#of coefficients at α = 0.01
coef_alfa1 <- summary(democracy_gdp_model)$coefficients[summary(democracy_gdp_model)$coefficients[, "Pr(>|t|)"] < 0.01, ]
coef_alfa1

# finding the outliers
democracy_gdp_model$residuals <- resid(democracy_gdp_model)
out_liers <-  boxplot.stats(democracy_gdp_model$residuals)["out"]

cat("The countries with outliers are:", paste0(components[democracy_gdp_model$residual %in% out_liers[[1]], "Country"], sep =","),"\n")

outlier_table <- outlier_table %>% mutate(residual = democracy_gdp_model$residual) %>% arrange(residual)

head(outlier_table,5)
tail(outlier_table,5)

```

Things that can affect GDP besides democracy can be natural resources for example (in countries like Qatar, United Arab Emirates and Luxembourg).
