---
title: "52414 - lab 1"
author: "52414"
date: "10/5/2023"
output: html_document
---

```{r}
library(tidyverse) # This includes dplyr, stringr, ggplot2, .. 
library(data.table)
library(rworldmap) # world map
library(ggthemes)
library(reshape2) # melt: change data-frame format long/wide
library(e1071) # skewness and kurtosis
library(rvest)
library(corrplot)
library(moments)
library(spatstat.geom)

```

**Solution:** (Fill code, text, plots etc.)

# 1.a. Loading the data via URL connection:

```{r, cache=TRUE}
democracy <- read_html("https://en.wikipedia.org/wiki/Democracy_Index")
all.tables = html_nodes(democracy, "table")  
regions <- as.data.frame(html_table(all.tables[4], fill = TRUE))
countries <- as.data.frame(html_table(all.tables[6], fill = TRUE))
components <- as.data.frame(html_table(all.tables[7], fill = TRUE))
```

# 1.b.

```{r, cache=TRUE}
countries_rate <- as.data.frame(countries)
top_countries <- countries_rate %>% select(Country, X2022) %>% arrange(desc(X2022)) %>% head(5)
bottom_countries <- countries_rate %>% select(Country, X2022) %>% arrange(X2022) %>% head(5)
Average <- rowMeans(countries_rate %>% select(-c(Region,X2022.rank,Country, Regime.type)))
top_countries_avg <- countries_rate %>% mutate(Average) %>% select(Country,Average) %>% arrange(desc(Average)) %>% head(5)
bottom_countries_avg <- countries_rate %>% mutate(Average) %>% select(Country, Average) %>% arrange(Average) %>% head(5)
```

YOUR ANALYSIS HERE:

Top countries in democracy index for 2022:

```{r}
top_countries
```

By 2006 -2022 Average:

```{r}
top_countries_avg
```

Lowest countries in democracy index for 2022:

```{r}
bottom_countries
```

By 2006 -2022 Average:

```{r}
bottom_countries_avg
```

# 2.a.

```{r, cache=TRUE}
p2 <- ggplot(countries)
p2 <- p2 + geom_boxplot(aes(x=Region, y=X2022)) +theme(text = element_text(size = 6))
p3 <- countries %>% select(Country,Region,X2022) %>% filter(Region == "Middle East and North Africa")
p4 <- countries %>% select(Country,Region,X2022) %>% filter(Region == "Western Europe")
out_3 <- boxplot.stats(p3$X2022)$out
out_ind_3 <- which(p3$X2022 %in% c(out_3))
out_4 <- boxplot.stats(p4$X2022)$out
out_ind_4 <- which(p4$X2022 %in% c(out_4))
p3[out_ind_3,]
p4[out_ind_4,]

```


2.b.

```{r, cache=TRUE}
p5 <- ggplot(countries,aes(X2022)) + geom_density(alpha=.2, fill= "#00BFC4") + facet_wrap(~Region)
p6 <- countries %>% 
     group_by(Region) %>%
     summarize(Mean = mean(X2022), Variance = var(X2022), Skewness = skewness(X2022), Kurtosis = kurtosis(X2022))
```

YOUR ANALYSIS HERE

3.a.

```{r, cache=TRUE}
Countries_graphs <- function(df, names){
  colnames <- colnames(df)
  filtered_df <- df %>% filter(Country %in% names | Region %in% names ) %>%
    select(colnames(countries)[which((colnames >= "X2006" & colnames <= "X2022") | (colnames == "Region" | colnames == "Country"))])
  
  
  
  
  
  
  
  
}
```

YOUR ANALYSIS HERE

3.b.

```{r, cache=TRUE}
# YOUR CODE HERE
```

YOUR ANALYSIS HERE

4.  

```{r, cache=TRUE}
# YOUR CODE HERE
```

YOUR ANALYSIS HERE

5.a.

```{r, cache=TRUE}
# YOUR CODE HERE
```

YOUR ANALYSIS HERE

5.b.

```{r, cache=TRUE}
# YOUR CODE HERE
```

YOUR ANALYSIS HERE

6.a.

```{r, cache=TRUE}
# YOUR CODE HERE
```

YOUR ANALYSIS HERE

6.b.

```{r, cache=TRUE}
# YOUR CODE HERE
```

YOUR ANALYSIS HERE

6.c.

```{r, cache=TRUE}
# YOUR CODE HERE
```

YOUR ANALYSIS HERE

7

```{r, cache=TRUE}
# YOUR CODE HERE
```

YOUR ANALYSIS HERE

8.a.

```{r, cache=TRUE}
# YOUR CODE HERE
```

YOUR ANALYSIS HERE

8.b.

```{r, cache=TRUE}
# YOUR CODE HERE
```

YOUR ANALYSIS HERE
